### 2024-01-04 analysis/RD design analysis.Rmd
Add files via upload

---
title: "RD design analysis"
author: "Xuyuan Zhang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
geometry: "left=1.5cm,right=1.5cm,top=1cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(fig.height=4)
knitr::opts_chunk$set(fig.width=7)
knitr::opts_chunk$set(warning=FALSE)
options(encoding = "UTF-8")
set.seed(123)
```

## structure

1. we first report the figure and export the figure to the folder named figures in directory
2. we then report the main result of the regressions
3. lastly, we report the placebo test and the robustness check

```{r}
# these are the packages that we need to import
library(haven)
library(ggplot2)
library(dplyr)
library(rdrobust)
library(plm)
library(pander) # this is for table showing
library(latex2exp) # this is to export it
```

```{r}
# need to place data on the base directory
data <- read_dta("../template.dta")
panel_data <- pdata.frame(data, index = c("id", "year"))
panel_data$L_jiadian <- lag(panel_data$jiadian, 1)
panel_data$L_kind <- lag(panel_data$kind, 1)
panel_data$L_hotel <- lag(panel_data$hotel, 1)
panel_data$L_shop_mall <- lag(panel_data$shop_mall, 1)
panel_data$L_museum <- lag(panel_data$museum, 1)
panel_data$L_old <- lag(panel_data$old, 1)
panel_data$L_ktv <- lag(panel_data$ktv, 1)
panel_data$L_mid <- lag(panel_data$mid, 1)
panel_data$L_prim <- lag(panel_data$prim, 1)
panel_data$L_west_food <- lag(panel_data$west_food, 1)
panel_data$L_super <- lag(panel_data$super, 1)
panel_data$L_sub <- lag(panel_data$sub, 1)
panel_data$L_park <- lag(panel_data$park, 1)
```

we use these control variables

```{r}
control_variables <- c(
  "jiadian", "kind", "hotel", "shop_mall", "museum", "old", "ktv", "mid", "prim", "west_food", "super", "sub", "park", 
  "area", "bedroom", "toilet", "house_age", "floor_level", "green_ratio", "total_building", "total_floor_number", "living_room", "elevator_ratio", "kitchen", "floor_ratio", "total_resident",
  "L_jiadian", "L_kind", "L_hotel", "L_shop_mall", "L_museum", "L_old", "L_ktv", "L_mid", "L_prim", "L_west_food", "L_super", "L_sub", "L_park",
  "pm25", "pop", "light",
  "non_online_effect", "ln_negotiation_period", "ln_watch_time", "ln_nego_changes",
  "density", "lianjia_5", "other_5", "ln_lead", "ln_watch_people"
)
```

```{r}
rdplot(y = filtered_data$ln_income, 
       x = filtered_data$density, 
       c = 0.2, p =1,  nbins = 20, 
       title = "RD-plot",
       x.label = "nearest distance to lianjia's stores",
       y.label = "lianjia's income in the community")
```

note that the figure only shows points that are within the bandwidth, the whole points are shown in Appendix of the paper.
```{r}
h = 0.151
cutoff = 0.5
running_var = panel_data$nearest_store_distances

filtered_data <- panel_data[abs(running_var - cutoff) <= h, ]

rdplot(y = filtered_data$ln_income, 
       x = filtered_data$nearest_store_distances, 
       c = cutoff, p = 1, nbins = 20, 
       title = "RD-plot",
       x.label = "nearest distance to lianjia's stores",
       y.label = "lianjia's income in the community")
```

```{r}
result <- rdrobust(y = panel_data$ln_income, x = panel_data$density, 
                   c = 0.2, covs = panel_data[, control_variables],
                   p = 1) # first using default clustering
```

```{r}
pander(cbind(result$coef, result$pv))
```

```{r}
result
```