### 2024-05-04 result_tables/summary_statistics.tex
Add files via upload

\begin{tabular}{p{5cm}lllll}
\toprule
Name & Mean without lianjia & SD without lianjia & Mean with lianjia & SD with lianjia & Difference \\
\midrule
The time it takes before a deal is made. & 13.59 & 14.56 & 17.09 & 17.50 & -3.501 (-49.823***) \\
price\_concession & -0.0367 & 0.0318 & -0.0351 & 0.0294 & -0.00200 (-12.088***) \\
percentage of lianjia to all brokerages & 0 & 0 & 0.206 & 0.166 & -0.206 (-384.429***) \\
number of brokerages within 410 meters, which is the cutoff of RD & 5.362 & 6.871 & 12.66 & 8.410 & -7.300 (-217.614***) \\
The number of people watching a listing. & 17.61 & 29.11 & 20.29 & 29.40 & -2.682 (-21.214***) \\
The final agreed price. & 260.5 & 254.9 & 351.9 & 292.1 & -91.48 (-76.662***) \\
without online platformization influence & 0.195 & 0.396 & 0.233 & 0.423 & -0.0380 (-21.384***) \\
The number of times a listing is watched. & 1121 & 1827 & 1233 & 1969 & -112.4 (-13.636***) \\
The number of times a negotiation was held. & 4.769 & 7.686 & 5.683 & 10.78 & -0.914 (-22.178***) \\
The period over which negotiations took place. & 150.5 & 187.5 & 166.6 & 239.1 & -16.06 (-17.075***) \\
Referring to electronic shops. & 1.489 & 3.566 & 1.942 & 4.365 & -0.453 (-26.004***) \\
Referring to proximity to kindergartens & 8.665 & 6.229 & 11.76 & 6.081 & -3.097 (-116.630***) \\
Referring to proximity to hotels & 3.211 & 5.287 & 5.428 & 6.315 & -2.217 (-87.264***) \\
Referring to shopping mall. & 4.554 & 7.539 & 6.698 & 8.489 & -2.144 (-61.405***) \\
Distance to the nearest museum. & 0.617 & 1.533 & 1.023 & 1.869 & -0.406 (-54.395***) \\
Referring to old care systems. & 0.894 & 1.628 & 1.339 & 1.950 & -0.446 (-56.870***) \\
Referring to KTV and some entertainment venues. & 5.179 & 7.853 & 7.305 & 7.759 & -2.126 (-63.096***) \\
Referring to middle schools. & 2.059 & 2.329 & 3.368 & 2.853 & -1.309 (-115.077***) \\
Referring to primary schools. & 2.812 & 2.846 & 4.343 & 3.205 & -1.531 (-116.172***) \\
Referring to the availability of western food nearby. & 3.880 & 7.709 & 7.317 & 10.02 & -3.437 (-87.756***) \\
Referring to proximity to supermarkets (measured by number within given distance & 3.155 & 3.374 & 4.499 & 3.686 & -1.344 (-87.586***) \\
Referring to proximity to subway stations. & 0.683 & 0.945 & 1.111 & 1.099 & -0.429 (-96.038***) \\
Referring to parks. & 3.422 & 4.575 & 4.569 & 3.944 & -1.147 (-62.708***) \\
The area of a property. & 90.82 & 48.57 & 84.97 & 39.15 & 5.845 (31.050***) \\
The number of bedrooms in a property. & 2.338 & 0.816 & 2.202 & 0.730 & 0.136 (40.958***) \\
The number of toilets in a property. & 1.306 & 0.579 & 1.246 & 0.455 & 0.0600 (26.915***) \\
The age of the house. & 18.08 & 11.68 & 20.73 & 11.77 & -2.651 (-52.316***) \\
The level on which a particular room or apartment is, within a building. & 1.854 & 0.975 & 1.933 & 0.931 & -0.0790 (-19.324***) \\
The ratio of the green space to the total plot area. & 0.309 & 0.237 & 0.300 & 0.107 & 0.00900 (12.288***) \\
The total number of buildings in an area. & 26.88 & 56.66 & 20.51 & 50.54 & 6.371 (27.653***) \\
The number of floors in a building. & 12.75 & 8.454 & 12.97 & 8.485 & -0.221 (-6.035***) \\
The number of living rooms in a property. & 1.445 & 0.504 & 1.341 & 0.492 & 0.104 (48.345***) \\
The ratio of elevators to the total number of floors. & 0.453 & 0.413 & 0.408 & 0.279 & 0.0450 (30.475***) \\
The number of kitchens in a property. & 0.983 & 0.150 & 0.983 & 0.125 & 0 (-0.459) \\
The ratio of the floor area to the total plot area. & 4.942 & 329.1 & 2.702 & 9.385 & 2.240 (2.367**) \\
The total number of residents in an area. & 995.8 & 1079 & 901.8 & 956.5 & 93.97 (21.484***) \\
Air quality measure. & 44.08 & 13.29 & 45.74 & 13.89 & -1.664 (-28.266***) \\
Population density. & 15506 & 16336 & 24634 & 18866 & -9100 (-118.769***) \\
Night time lights. & 32.62 & 13.59 & 38.41 & 11.92 & -5.782 (-105.485***) \\
\bottomrule
\end{tabular}