### 2024-05-04 analysis/RD design for density.Rmd
Add files via upload

---
title: "RD design analysis"
author: "Xuyuan Zhang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
geometry: "left=1.5cm,right=1.5cm,top=1cm,bottom=1.5cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(fig.height=4)
knitr::opts_chunk$set(fig.width=7)
knitr::opts_chunk$set(warning=FALSE)
options(encoding = "UTF-8")
set.seed(123)
```

## structure

1. we first report the figure and export the figure to the folder named figures in directory
2. we then report the main result of the regressions
3. lastly, we report the placebo test and the robustness check

```{r}
# these are the packages that we need to import
library(haven)
library(ggplot2)
library(dplyr)
library(rdrobust)
library(plm)
library(pander) # this is for table showing
library(latex2exp) # this is to export it
```

```{r}
# need to place data on the base directory
data <- read_dta("../template.dta")
panel_data <- pdata.frame(data, index = c("id", "year"))
panel_data$L_jiadian <- lag(panel_data$jiadian, 1)
panel_data$L_kind <- lag(panel_data$kind, 1)
panel_data$L_hotel <- lag(panel_data$hotel, 1)
panel_data$L_shop_mall <- lag(panel_data$shop_mall, 1)
panel_data$L_museum <- lag(panel_data$museum, 1)
panel_data$L_old <- lag(panel_data$old, 1)
panel_data$L_ktv <- lag(panel_data$ktv, 1)
panel_data$L_mid <- lag(panel_data$mid, 1)
panel_data$L_prim <- lag(panel_data$prim, 1)
panel_data$L_west_food <- lag(panel_data$west_food, 1)
panel_data$L_super <- lag(panel_data$super, 1)
panel_data$L_sub <- lag(panel_data$sub, 1)
panel_data$L_park <- lag(panel_data$park, 1)
```

we use these control variables

```{r}
control_variables <- c(
  "jiadian", "kind", "hotel", "shop_mall", "museum", "old", "ktv", "mid", "prim", "west_food", "super", "sub", "park", 
  "area", "bedroom", "toilet", "house_age", "floor_level", "green_ratio", "total_building", "total_floor_number", "living_room", "elevator_ratio", "kitchen", "floor_ratio", "total_resident",
  "pm25", "pop", "light",
  "non_online_effect", "ln_negotiation_period", "ln_watch_time", "ln_nego_changes", "ln_lead", "ln_watch_people"
)
testing_variables <- c("ln_negotiation_period", "ln_watch_time", "ln_nego_changes", "ln_lead", "ln_watch_people", "ln_end_price")
```

```{r}
result <- rdrobust(y = panel_data$ln_income, x = panel_data$density, 
                   c = 0.2,
                   p = 1) # first using default clustering
```

```{r}
result
```

```{r}
pander(cbind(result$coef, result$pv))
```

```{r}
result <- rdrobust(y = panel_data$ln_income, x = panel_data$density, 
                   c = 0.41,
                   p = 1) # first using default clustering
```

```{r}
result
```

```{r}
pander(cbind(result$coef, result$pv))
```

```{r}
result
```

note that the figure only shows points that are within the bandwidth, the whole points are shown in Appendix of the paper.

```{r}
h = 0.095
cutoff = 0.5
running_var = panel_data$density

filtered_data <- panel_data[abs(running_var - cutoff) <= h, ]

par(mfrow=c(3,1))
rdplot(y = filtered_data$ln_income, 
       x = filtered_data$density, 
       c = cutoff, nbins = 40, p = 1,
       title = "RD-plot",
       x.label = "nearest distance to lianjia's stores",
       y.label = "lianjia's income in the community")

rdplot(y = filtered_data$ln_income, 
       x = filtered_data$density, 
       c = cutoff, nbins = 40, p = 2,
       title = "RD-plot",
       x.label = "nearest distance to lianjia's stores",
       y.label = "lianjia's income in the community")

rdplot(y = filtered_data$ln_income, 
       x = filtered_data$density, 
       c = cutoff, nbins = 40, p = 3,
       title = "RD-plot",
       x.label = "nearest distance to lianjia's stores",
       y.label = "lianjia's income in the community")
```

checkings:

1. Donut Hole: Exclude observations close to the cutoff and estimate the effect with the remaining data. If the discontinuity persists despite the absence of data around the cutoff, this provides evidence against a model misspecification.
2. Mccrary (2008) test: Conduct a McCrary density test to check if there is a discontinuity in the density of the running variable at the cutoff. A significant discontinuity may suggest manipulation or sorting around the cutoff.
3. Placebo test: If you have pre-treatment covariates, you can test whether these covariates also jump at the cutoff. If they do, this might suggest manipulation or a violation of the RD design assumptions. If they don’t, it strengthens the case for a true treatment effect.

```{r}
cutoff
donut_width <- 0.011
panel_data_donut <- panel_data[!(panel_data$nearest_store_distances >= (cutoff - donut_width) & 
                                 panel_data$nearest_store_distances <= (cutoff + donut_width)),]
result_donut <- rdrobust(y = panel_data_donut$ln_income, 
                         x = panel_data_donut$nearest_store_distances, 
                         c = cutoff,
                        p = 1, h = 0.066, b = 0.135)
pander(cbind(result_donut$coef, result_donut$pv))
print(result_donut)
```

```{r}
donut_width <- 0.015
panel_data_donut <- panel_data[!(panel_data$nearest_store_distances >= (cutoff - donut_width) & 
                                 panel_data$nearest_store_distances <= (cutoff + donut_width)),]
result_donut <- rdrobust(y = panel_data_donut$ln_income, 
                         x = panel_data_donut$nearest_store_distances, 
                         c = cutoff,  covs = panel_data_donut[, control_variables],
                        p = 1, h = 0.095, b = 0.185)
pander(cbind(result_donut$coef, result_donut$pv))
print(result_donut)
```

```{r}

library(rdd)

# Give it the running variable and the cutpoint
# it will automatically produce a plot and select the number of bins and the bandwidth
# The output will be the p-value for the presence of a discontinuity
DCdensity(panel_data$nearest_store_distances, cutpoint = 0.41, bw = 0.066, verbose = TRUE, plot = FALSE)
```

```{r}
library(rdd)

# Give it the running variable and the cutpoint
# it will automatically produce a plot and select the number of bins and the bandwidth
# The output will be the p-value for the presence of a discontinuity
DCdensity(panel_data$nearest_store_distances, cutpoint = 0.41, bw = 0.095, verbose = TRUE, plot = FALSE)
```

Conducting placebo test. We shall using the same validation bandwidth and bias correction bandwidth. The results are good overall

```{r}
for (validation in testing_variables) {
  rd_result <- rdrobust(y = panel_data[[validation]], x = panel_data$nearest_store_distances
                        , c = 0.41, h = 0.066, b = 0.135)
  cat("Covariate:", validation, "\n")
  print(summary(rd_result))
  cat("\n\n")

}
```