### 2024-06-23 result_tables/code_book_table.tex
modified:   analysis/final_analysis.do 	modified:   analysis/individual.do 	modified:   cleaning/clean district.do 	modified:   manuscript/main.aux 	modified:   manuscript/main.bbl 	modified:   manuscript/main.blg 	modified:   manuscript/main.fdb_latexmk 	modified:   manuscript/main.fls 	modified:   manuscript/main.log 	new file:   manuscript/main.pdf 	deleted:    manuscript/main.synctex(busy) 	modified:   manuscript/main.synctex.gz 	modified:   manuscript/main.tex 	new file:   result_tables/code_book_table.tex 	modified:   result_tables/formulate_the_tables.ipynb 	new file:   result_tables/st

\begin{tabular}{lll}
\toprule
Name & Label & Dimension \\
\midrule
income & The income lianjia in this given district/housing. & $10^5 \times$ \textyen \\
lead\_times & The time it takes before a deal is made. & counts \\
price\_concession & price changes (ending price - starting price) / starting price & \% \\
density & percentage of lianjia to all brokerages & \% / 100 \\
broker\_410 & number of other brokerages within 410 meters, which is the cutoff of RD & counts \\
watching\_people & The number of people watching this listing. & counts \\
end\_price & The final agreed price. & \textyen \\
non\_online\_effect & without online platformization influence & bool indicator \\
watched\_times & The number of times a listing is watched. & counts \\
nego\_times & The number of times a negotiation was held. & counts \\
nego\_period & The period over which negotiations took place. & days \\
jiadian & Referring to electronic shops. & counts \\
kind & Referring to proximity to kindergartens & counts \\
hotel & Referring to proximity to hotels & counts \\
shop\_mall & Referring to shopping mall. & counts \\
museum & Distance to the nearest museum. & counts \\
old & Referring to old care systems. & counts \\
ktv & Referring to KTV and some entertainment venues. & counts \\
mid & Referring to middle schools. & counts \\
prim & Referring to primary schools. & counts \\
west\_food & Referring to the availability of western food nearby. & counts \\
super & Referring to proximity to supermarkets (measured by number within given distance & counts \\
sub & Referring to proximity to subway stations. & counts \\
park & Referring to parks. & counts \\
area & The area of a property. & $m^2$ \\
bedroom & The number of bedrooms in a property. & counts \\
toilet & The number of toilets in a property. & counts \\
house\_age & The age of the house. & years \\
floor\_level & The level on which a particular room or apartment is, within a building. & categories \\
green\_ratio & The ratio of the green space to the total plot area. & \% / 100 \\
total\_building & The total number of buildings in an area. & counts \\
total\_floor\_number & The number of floors in a building. & counts \\
living\_room & The number of living rooms in a property. & counts \\
elevator\_ratio & The ratio of elevators to the total number of floors. & \% / 100 \\
kitchen & The number of kitchens in a property. & counts \\
floor\_ratio & The ratio of the floor area to the total plot area. & fraction \\
total\_resident & The total number of residents in an area. & counts \\
pm25 & Air quality measure. & mass/volume \\
pop & Population density. & people/$km^2$ \\
light & Night time lights. & lux \\
\bottomrule
\end{tabular}