### 2024-07-09 result_tables/rdd_result_summary_robust.tex

% latex table generated in R 4.2.0 by xtable 1.8-4 package
% Mon Jul  8 21:13:07 2024
\begin{table}[ht]
\centering
\begin{tabular}{lllllllr}
  \hline
Method & Kernel & Estimate & SE & Z & PValue & Bandwidth & EffectiveObs \\ 
  \hline
mserd & uniform & -0.0584 & 0.0263 & -2.22 & 0.0263 & 0.0773 & 37645 \\ 
  mserd & triangular & -0.0697 & 0.0314 & -2.22 & 0.0267 & 0.066 & 32020 \\ 
  cerrd & uniform & -0.0874 & 0.0361 & -2.42 & 0.0156 & 0.0418 & 19988 \\ 
  cerrd & triangular & -0.0663 & 0.0434 & -1.53 & 0.126 & 0.0357 & 17191 \\ 
   \hline
\end{tabular}
\caption{RD Estimates with Different Bandwidth Selection Methods and Kernels (robust method with log(income))} 
\label{tab:rd_bandwidth_kernel_results_robust}
\end{table}