### 2024-09-06 figures/distribution_of_hp_and_broker.tex
Update distribution_of_hp_and_broker.tex

\begin{figure}[h!]
 \centering
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Beijing.pdf}
    \end{minipage}
    \hfill
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Chengdu.pdf}
    \end{minipage}
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Chongqing.pdf}
    \end{minipage}

    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Guangzhou.pdf}
    \end{minipage}
    \hfill
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Hangzhou.pdf}
    \end{minipage}
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Nanjing.pdf}
    \end{minipage}

    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Shanghai.pdf}
    \end{minipage}
    \hfill
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Shenzhen.pdf}
    \end{minipage}
    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Tianjin.pdf}
    \end{minipage}

    \begin{minipage}{0.328\textwidth}
        \includegraphics[width=\linewidth]{figures/distribution_of_hp_and_broker/Wuhan.pdf}
    \end{minipage}
    \hfill

\caption{The distribution of housing prices and brokers in different cities}
\label{fig:distribution_of_housing_price_brokers_in_different_cities}

Note: The heat map represents the distribution of housing prices and the standard ellipse represents the distribution of brokers, which is calculated by the mean and standard deviation of the latitude and longitude of brokers. The larger the ellipse, the more sparse the brokerages' distribution is. The Base Map is from Google Map.
 \end{figure}